A Tale of Two Transects

### 2020-02-12 writing/transect_paper.Rmd
clarify methods and results

loadd(
  reads_table,
  region_table,
  asv_table,
  otu_table,
  demuxlength,
  demuxqual,
  readcounts,
  rawcounts,
  demuxcounts,
  regioncounts,
  cache = cache
)

@@

Ectomycorrhizas (ECM) are a symbiosis between fungi and plant root tips.
The fungi provide nutrients and protection from pathogens in exchange for sugards from the plant [@Smith2010].
Approximately 8% of described fungal species [@ainsworth2008ainsworth;@Rinaldi2008] and 2% of land plant species [@Brundrett2017a] form ECM.
The plants include ecologically and economically important stand-forming trees belonging to both temperate/boreal groups such as Pinaceae and Fagaceae, and tropical groups such as Dipterocarpaceae, and Fabaceae tr. Amherstieae.

@@

Turnover scale is the distance at which two communities can be considered to be independent samples of the local species pool.
Knowledge of turnover scale is import when planning studies of local diversity and its environmental correlates.
It varies between different ecosystems and taxonomic groups.
Turnover scale is often measured by the range at which Moran's I statistic indicates significant autocorrelation, or by fitting a function to an empirical distance-decay curve community dissimilarity vs. distance [@legendre2012].

@@

We attempted to denoise both long and short PacBio amplicons using `DADA2` version `r packageVersion("dada2")` [@callahan2016; @callahan2019] according to the steps outlined in  the supplementary information in @callahan2019.
However, only `r readd(region_table)$ASVs[readd(region_table)$region=="long"]` amplicon sequence variants (ASVs) were obtained for the long amplicons, representing `r percent(readd(region_table)$reads[readd(region_table)$region == "long"]/as.integer(gsub(" ", "", readd(demuxcounts)["pb_500"])), format = "d")` of the trimmed reads.
We believe that this poor performance was due to a combination of long read length and low sequencing depth relative to community diversity.
The `DADA2` algorithm requires that the seed sequence of each ASV be represented by at least two error-free reads [@callahan2016].
If sequencing errors are uniformly distributed, then the probability that a given read will be error-free is $(1-\epsilon)^L$, where $\epsilon$ is the sequencing error rate and $L$ is the read length in base pairs.
Then the number of reads of a given sequence that would be required to obtain two error-free reads in expectation is $2/(1-\epsilon)^L$.
For the combination of long reads (median $L=`r demuxlength['pb_500']` \text{ bp}$ after trimming) and moderate error rate (mean $\epsilon= `r format(demuxqual['pb_500'], digits = 4)`$ based on ccs quality scores) for the long amplicon in this study, the expected number of reads required to achieve two error-free reads is `r format(as.integer(round(2/(1 - demuxqual['pb_500'])^demuxlength['pb_500'])), big.mark = " ")`.
Given the high diversity relative to sequencing depth in this study (`r region_table$ASVs[region_table$seq_run == "pb_483" & region_table$region == "short"]` ASVs based on PacBio short amplicons, `r demuxcounts["pb_500"]` trimmed long amplicon reads), this requirement could not have been met for the long amplicons except by the most abundant sequences.
In comparison, the equivalent requirement for the short amplicon ($L=`r demuxlength['pb_483']` \text{ bp}$, $\epsilon=`r demuxqual['pb_483']`$) is only `r round(2/(1 - demuxqual['pb_483'])^demuxlength['pb_483'], 1)` reads.
We therefore developed a new workflow to assemble ASVs from the long amplicons, as follows:

@@

Additionally, empirical variograms were generated by plotting mean community dissimilarity as a function of spatial distance and temporal separation, and fit using a joint exponential spatio-temporal model using the `fit.StVariogram` function from `gstat` version `r packageVersion("gstat")` [@R-gstat].
The mean dissimilarity between samples taken in different transects (distance $\approx$ 30 km) was also included as a single point for each year for the purposes of fitting the variogram function.

@@

The number of sequencing reads and ASVs at each stage in the bioinformatics pipeline are shown in Table \@ref(tab:bioinfo).
Long (short) amplicon sequencing with PacBio yielded `r rawcounts["pb_500"]` (`r rawcounts["pb_483"]`) circular consensus reads, of which `r demuxcounts["pb_500"]` (`r demuxcounts["pb_483"]`) were successfully demultiplexed.
The 5.8S region was detected in `r regioncounts["pb_500"]` (`r regioncounts["pb_483"]`) reads.
Ion Torrent sequencing yielded `r demuxcounts["is_057"]` demultiplexed reads, with the 5.8S region detected in `r regioncounts["is_057"]`.

@@

@@

readd(vennplot_OTU) %>% {
  plot(
    select(., set, OTUs) %>% deframe() %>% replace_na(0) %>% eulerr::venn(),
    legend = FALSE,
    labels = FALSE,
    quantities = .$label,
    main = "OTU"
  )
}

@@

@@

@@

```{r variofit, results="asis"}
variog_data %>%
  select(mantel_0, mantel_1, variofit_0 = variofit2, variofit_1 = variofitST2, metric, amplicon, tech, guild, algorithm) %>%
  pivot_longer(1:4, names_to = c(".value", "timelag"), names_sep = "_") %>%
  mutate(params = map(variofit, broom::tidy)) %>%
  unnest(params) %>%
  bind_cols(
    pmap_dfr(
      .,
      function(mantel, variofit, term, ...) {
        if (mantel) {
          tryCatch(broom::confint_tidy(object = variofit, parm = term) %>% as.list %>% inset2("term", term) %>% as.data.frame,
                   error = function(e) tibble(term = term, conf.low = NA_real_, conf.high = NA_real_))
        } else {
          tibble(term = term, conf.low = NA_real_, conf.high = NA_real_)
        }
      }) %>%
      select(-term)
  ) %>%
  filter(term %in% c("range", "timerange")) %>%
  mutate_at("conf.low", replace_na, 0) %>%
  mutate(value = ifelse(
    is.na(estimate) | is.na(conf.high),
    "-",
    sprintf("%.2g (%.2g–%.2g)", estimate, conf.low, conf.high)
  )) %>%
  filter(timelag == "0" | term == "timerange") %>%
  pivot_wider(names_from = "term", values_from = "value", id_cols = c("guild", "amplicon", "tech", "metric", "algorithm")) %>%
  select(Guild = guild, Amplicon = amplicon, Tech = tech, Algorithm = algorithm, Metric = metric, "Space Range (m)" = range, "Time Range (a)" = timerange) %>%
  arrange(Guild, Amplicon, Tech, Algorithm, Metric) %>%
  kable(booktabs = TRUE, caption = "(ref:variofit)", linesep = c(rep("", 5), "\\addlinespace"))
```

@@

```{r phylotax, fig.height = 2, fig.width = 3, fig.cap = "(ref:phylotax)"}
extree <-
  treeio::as.treedata(ape::read.tree(text = "(A:1,((B:1,C:1):1,((E:1,F:1):1,D:1):1):1);")) %>%
  tidytree::full_join(
    tibble(label = LETTERS[1:6],
           tax1 = c("unk", "Tax1", "Tax2", "Tax2", "unk", "unk"),
           tax2 = c("unk", "Tax2", "Tax2", "Tax1", "unk", "Tax1")),
    by = "label"
  )
ggtree::ggtree(extree) +
  ggtree::geom_tiplab(x = 4, align = 2) +
  xlim(0, 6) +
  ggtree::geom_tiplab(
    aes(color = tax1, label = tax1),
    x = 4.7,
    hjust = 0.5
  ) +
  ggtree::geom_tiplab(
    aes(color = tax2, label = tax2),
    x = 5.7,
    hjust = 0.5
  ) +
  ggtree::geom_tiplab(x = 5.2, label = "/", hjust = 0.5) +
  ggtree::geom_nodelab(aes(label = node - 6), geom = "label") +
  ggtree::geom_hilight(node = 10, fill = "steelblue", extendto = 4) +
  ggtree::geom_hilight(node = 9, fill = "tomato", extendto = 4) +
  scale_color_manual(values = c(Tax1 = "steelblue", Tax2 = "tomato", unk = "grey50")) +
  scale_y_reverse()
```

(ref:full-length) Comparison of the length of sequences from different length amplicons and sequencing technologies.  Length distribution (*a*, *c*) and empirical cumulative distribution function (*b*, *d*) for short (*a*, *b*) and long (*c*, *d*) amplicons, respectively.

```{r full-length, fig.cap = "(ref:full-length)", fig.height = 6}
full_dens <- readd(reads_table) %>%
  dplyr::left_join(dplyr::select(datasets, seq_run, tech, amplicon)) %>%
  dplyr::filter(region %in% c("long", "short")) %>%
  dplyr::group_by(seq_run) %>%
  dplyr::mutate_at("amplicon", factor, levels = c("Short", "Long")) %>%
  dplyr::mutate_at("tech", factor, levels = c("PacBio", "Illumina", "Ion Torrent")) %>%
  dplyr::mutate(
    reads = reads / sum(reads)
  )

full_dens_short <- full_dens %>%
  filter(amplicon == "Short") %>%
  ggplot(aes(length, weight = reads, color = tech, group = tech)) +
  stat_density(bw = 0.5, geom = "line", position = "identity") +
  scale_color_brewer(type = "qual", guide = "none", palette = "Set2") +
  scale_y_continuous(name = "Density") +
  scale_x_continuous(name = NULL, labels = NULL, breaks = NULL) +
  ggnomics::facet_nested(rows = vars(tech), scales = "free_y") +
  theme(strip.background = element_blank())

full_dens_long <- full_dens %>%
  filter(amplicon == "Long") %>%
  ggplot(aes(length, weight = reads, color = tech, group = tech)) +
  stat_density(bw = 0.5, geom = "line", position = "identity") +
  scale_color_brewer(type = "qual", guide = "none", palette = "Set2") +
  scale_y_continuous(name = "Density") +
  scale_x_continuous(name = NULL, labels = NULL, breaks = NULL) +
  ggnomics::facet_nested(rows = vars(tech), scales = "free_y") +
  theme(strip.background = element_blank())

full_ecdf <- readd(reads_table) %>%
  dplyr::filter(region  %in% c("long", "short")) %>%
  left_join(select(datasets, seq_run, amplicon, tech, machine), by = "seq_run") %>%
  mutate_at("tech", factor, levels = c("PacBio", "Illumina", "Ion Torrent")) %>%
  dplyr::group_by(seq_run) %>%
  dplyr::mutate(reads = reads / sum(reads))

full_ecdf_short <- full_ecdf %>%
  filter(amplicon == "Short") %>%
  ggplot(aes(length, weight = reads, color = tech, group = tech)) +
  stat_ecdf(geom = "line") +
  scale_color_brewer(type = "qual",
                     name = NULL, palette = "Set2") +
  scale_y_continuous(name = "ECDF") +
  scale_x_continuous(name = NULL) +
  facet_grid( amplicon ~ ., scales = "free_x") +
  theme(
    legend.position = c(1, 0.02),
    legend.justification = c(1, 0),
    legend.background = element_blank(),
    strip.background = element_blank()
  )

full_ecdf_long <- full_ecdf %>%
  filter(amplicon == "Long") %>%
  ggplot(aes(length, weight = reads, color = tech, group = tech)) +
  stat_ecdf(geom = "line") +
  scale_color_brewer(type = "qual",
                     name = NULL, palette = "Set2", direction = -1) +
  scale_y_continuous(name = "ECDF") +
  scale_x_continuous(name = "Length (bp)") +
  facet_grid( amplicon ~ ., scales = "free_x") +
  theme(
    legend.position = c(1, 0.02),
    legend.justification = c(1, 0),
    legend.background = element_blank(),
    strip.background = element_blank()
  )

ggpubr::ggarrange(
  full_dens_short,
  full_ecdf_short,
  full_dens_long,
  full_ecdf_long,
  ncol = 1, heights = c(3, 1.7, 1.2, 2), labels = "auto", vjust = 1

)
```

@@

(ref:region-asv) Plot of total number of ASVs recovered vs. fraction of reads successfully mapped to ASVs for different rDNA regions extracted from long PacBio amplicons using `LSUx`. Some reggions were formed by concatenation of extracted regions, as follows: LSU = LSU1--D1--LSU2--D2--LSU3--D3--LSU4; ITS = ITS1--5.8S--ITS2; 32S = 5.8S--ITS2--LSU; long = ITS--LSU. For this figure, subregions were concatenated prior to generating ASVs with `DADA2`.

```{r regions, fig.cap="(ref:region-asv)"}
readd(region_table) %>%
  dplyr::filter(seq_run == "pb_500") %>%
  ggplot(aes(reads, ASVs, color = length_med, label = region)) +
  geom_point() +
  ggrepel::geom_text_repel() +
  scale_color_gradientn(colors = c("blue", "red", "green1"), limits = c(30, 2000), breaks = c(50, 150, 400, 1000), trans = "log10", name = "median length") +
  xlab("Mapped reads")
```

@@

```{r region-lengths, fig.cap = "(ref:region-length)", fig.height = 6}
baseregions <- c("ITS1", "5_8S", "ITS2", "LSU1", "D1", "LSU2", "D2", "LSU3", "D3",
               "LSU4")

region_limits <- dplyr::filter(regions, region %in% baseregions) %>%
  dplyr::mutate_at(
    "region",
    factor,
    levels = baseregions
    ) %>%
  tidyr::pivot_longer(
    cols = c("min_length", "max_length"),
    names_to = "limit",
    names_pattern = "(min|max)_length",
    values_to = "length"
  )

readd(reads_table) %>%
  dplyr::filter(region %in% baseregions,
         seq_run == "pb_500") %>%
  dplyr::mutate_at(
    "region",
    factor,
    levels = baseregions
    ) %>%
  ggplot(aes(length, weight = reads)) +
  stat_density(bw = 0.5, color = "#8da0cb", geom = "line") +
  geom_vline(data = region_limits,
             aes(xintercept = length),
             linetype = "dashed", alpha = 0.5) +
  facet_wrap(~region,ncol = 1, strip.position = "right", scales = "free_y") +
  scale_y_continuous(breaks = NULL, labels = NULL, name = "Read density") +
  scale_x_continuous(limits = c(0, 500), name = "Length (bp)")
```

@@

```{r ITS2-length, fig.cap = "(ref:ITS2-length)", fig.height = 6}
its2 <- readd(reads_table) %>%
  dplyr::left_join(dplyr::select(datasets, seq_run, tech, amplicon), by = "seq_run") %>%
  dplyr::mutate_at("tech", factor, levels = c("PacBio", "Illumina", "Ion Torrent")) %>%
  dplyr::filter(region == "ITS2") %>%
  arrange(tech, amplicon) %>%
  mutate(
    group = paste(tech, "-", amplicon) %>% factor(levels = unique(.))
  ) %>%
  dplyr::group_by(seq_run) %>%
  dplyr::mutate(
    reads = reads / sum(reads)
  )

its2_dens <- its2 %>%
  ggplot(aes(length, weight = reads, color = group, group = group)) +
  stat_density(bw = 0.5, geom = "line", position = "identity") +
  scale_color_brewer(type = "qual", guide = "none", palette = "Set2") +
  scale_y_continuous(name = "Density") +
  scale_x_continuous(name = NULL, labels = NULL, breaks = NULL) +
  ggnomics::facet_nested(rows = vars(tech, amplicon), scales = "free_y",
                         nest_line = TRUE) +
  theme(strip.background = element_blank())

its2_ecdf <- its2 %>%
  mutate(tech = "", amplicon = "") %>%
  ggplot(aes(length, weight = reads, color = group, group = group)) +
  stat_ecdf(geom = "line") +
  scale_color_brewer(type = "qual", name = NULL, palette = "Set2") +
  scale_y_continuous(name = "ECDF") +
  scale_x_continuous(name = "ITS2 Length (bp)") +
  ggnomics::facet_nested(rows = vars(tech, amplicon), scales = "free_y",
                         nest_line = TRUE) +
  theme(
    strip.background = element_blank(),
    legend.position = c(0.98, 0.02),
    legend.justification = c(1, 0)
  )

ggpubr::ggarrange(its2_dens, its2_ecdf, ncol = 1, labels = "auto", heights = c(4, 2))
```

@@

```{r short-its, fig.cap = "(ref:short-its)"}
readd(big_seq_table_ITS2) %>%
  colSums() %>%
  enframe(name = "ITS2", value = "nreads") %>%
  mutate(length = nchar(ITS2)) %>%
  filter(length <= 140) %>%
  mutate_at("ITS2", ~tzara::seqhash(chartr("T", "U", .))) %>%
  semi_join(readd(taxon_reads), ., by = c("label" = "ITS2")) %>%
  # group_by(ITS2) %>%
  # filter(!"Short" %in% amplicon) %>%
  taxon_plot(kingdom, y = reads)
```

@@

```{r ECM, fig.cap = "(ref:ecm)"}
  readd(taxon_reads) %>%
  mutate_at("reference", replace_na, "All") %>%
  taxon_plot(ECM, kingdom == "Fungi") +
  theme(legend.position = "bottom", legend.title = element_blank())
```

>>>>

@@

Ectomycorrhizas (ECM) are a symbiosis between fungi and plant root tips.
The fungi provide nutrients and protection from pathogens in exchange for sugards from the plant [@Smith2010].
Approximately 8% of described fungal species [@ainsworth2008ainsworth;@Rinaldi2008] and 2% of land plant species [@brundrett2017a] form ECM.
The plants include ecologically and economically important stand-forming trees belonging to both temperate/boreal groups such as Pinaceae and Fagaceae, and tropical groups such as Dipterocarpaceae, and Fabaceae tr. Amherstieae.

@@

Turnover scale is the distance at which two communities can be considered to be independent samples of the local species pool.
Knowledge of turnover scale is import when planning studies of local diversity and its environmental correlates.
It varies between different ecosystems and taxonomic groups.
Turnover scale is often measured by the range at which a Mantel correlogram indicates significant autocorrelation, or by fitting a function to an empirical distance-decay curve community dissimilarity vs. distance [@legendre2012].

@@

We attempted to denoise both long and short PacBio amplicons using `DADA2` version `r packageVersion("dada2")` [@callahan2016; @callahan2019] according to the steps outlined in  the supplementary information in @callahan2019.
However, only `r readd(region_table)$ASVs[readd(region_table)$region=="long"]` amplicon sequence variants (ASVs) were obtained for the long amplicons, representing `r percent(readd(region_table)$reads[readd(region_table)$region == "long"]/as.integer(gsub(" ", "", readd(demuxcounts)["pb_500"])), format = "d")` of the trimmed reads.
We believe that this poor performance was due to a combination of long read length and low sequencing depth relative to community diversity.
The `DADA2` algorithm requires that the seed sequence of each ASV be represented by at least two error-free reads [@callahan2016].
If sequencing errors are uniformly distributed, then the probability that a given read will be error-free is $(1-\epsilon)^L$, where $\epsilon$ is the sequencing error rate and $L$ is the read length in base pairs.
Then the number of reads of a given sequence that would be required to obtain two error-free reads in expectation is $2/(1-\epsilon)^L$.
For the combination of long reads (median $L=`r readd(demuxlength)['pb_500']` \text{ bp}$ after trimming) and moderate error rate (mean $\epsilon= `r format(readd(demuxqual)['pb_500'], digits = 4)`$ based on ccs quality scores) for the long amplicon in this study, the expected number of reads required to achieve two error-free reads is `r format(as.integer(round(2/(1 - readd(demuxqual)['pb_500'])^readd(demuxlength)['pb_500'])), big.mark = " ")`.
Given the high diversity relative to sequencing depth in this study (`r readd(region_table)$ASVs[readd(region_table)$seq_run == "pb_483" & readd(region_table)$region == "short"]` ASVs based on PacBio short amplicons, `r readd(demuxcounts)["pb_500"]` trimmed long amplicon reads), this requirement could not have been met for the long amplicons except by the most abundant sequences.
In comparison, the equivalent requirement for the short amplicon ($L=`r readd(demuxlength)['pb_483']` \text{ bp}$, $\epsilon=`r readd(demuxqual)['pb_483']`$) is only `r round(2/(1 - readd(demuxqual)['pb_483'])^readd(demuxlength)['pb_483'], 1)` reads.
We therefore developed a new workflow to assemble ASVs from the long amplicons, as follows:

@@

Additionally, empirical distance-decay curves were generated by plotting mean community dissimilarity as a function of spatial distance, and fit to an exponential model of the form given by @Legendre2012 using the `nls` function in R.
Points in the empirical distance-decay curve were weighted by the number of comparisons within the distance class and the inverse of the distance for the purposes of model fitting.
For datasets where the Mantel corellogram indicated spatial correlation between samples taken in seperate years, the model was re-fit with an additional term to represent temporal correlation:
$$D = C_0 + C_1\left[1 - \exp\left(-3  \left(\frac{d}{a_d} + \frac{t}{a_t}\right)\right)\right]$$
where $D$, $d$, and $t$ represent the community dissimilarity, spatial distance, and time lag between samples, respectively, and the parameters are $C_0$, the community dissimilarity from replicate samples ("nugget"); $C_0 + C_1$, the community dissimilarity at long distances ("sill"); $a_d$ the spatial range at which the community dissimilarity has moved 95% of the way from "nugget" to "sill", and $a_t$ the equivalent temporal range.
95% confidence intervals were calculated for the spatial and temporal range parameters by profiling using the `MASS` package in R.

@@

The number of sequencing reads and ASVs at each stage in the bioinformatics pipeline are shown in Table \@ref(tab:bioinfo).
Long (short) amplicon sequencing with PacBio yielded `r readd(rawcounts)["pb_500"]` (`r readd(rawcounts)["pb_483"]`) circular consensus reads, of which `r readd(demuxcounts)["pb_500"]` (`r readd(demuxcounts)["pb_483"]`) were successfully demultiplexed.
The 5.8S region was detected in `r readd(regioncounts)["pb_500"]` (`r readd(regioncounts)["pb_483"]`) reads.
Ion Torrent sequencing yielded `r readd(demuxcounts)["is_057"]` demultiplexed reads, with the 5.8S region detected in `r readd(regioncounts)["is_057"]`.

@@

- Fraction where 5.8S was detected
- Length/ASV count/ASV fraction of extracted regions.

@@

readd(vennplot_OTU) %>% {
  plot(
    select(., set, OTUs) %>% deframe() %>% replace_na(0) %>% eulerr::venn(),
    legend = FALSE,
    labels = FALSE,
    quantities = .$label,
    main = "OTU"
  )
}
```

(ref:venn2) (*alternate to \@ref(fig:venn)*) Venn diagrams showing shared ITS2-based ASVs (left) and 97% OTUs (right) between different sequencing technologies from the same short amplicon library.
In each region, the number of ASVs/OTUs is given above, while the fractions of reads for each sequencing strategy are shown below.

```{r venn2, fig.cap = "(ref:venn2)", out.height = "30%", fig.show="hold"}
venndata(readd(asv_table), ASVs, cols = c("pb_483", "SH-2257", "is_057")) %>%
  vennplot_data(ASVs) %>% {
  plot(
     select(., set, ASVs) %>% deframe() %>% replace_na(0) %>% eulerr::venn(),
     legend = list(
         labels = c("PacBio Short", "Illumina", "Ion Torrent")
     ),
     quantities = .$label,
     main = "ASV",
     fills = c("grey85", "lightblue", "lightcoral"))
  }

venndata(readd(otu_table), OTUs, cols = c("pb_483", "SH-2257", "is_057")) %>%
  vennplot_data(OTUs) %>% {
  plot(
    select(., set, OTUs) %>% deframe() %>% replace_na(0) %>% eulerr::venn(),
    legend = FALSE,
    labels = FALSE,
    quantities = .$label,
    main = "OTU",
     fills = c("grey85", "lightblue", "lightcoral")
  )
}
```

(ref:venn3) (*alternate to \@ref(fig:venn)*) Venn diagrams showing shared ITS2-based ASVs (left) and 97% OTUs (right) between different amplicons.
In each region, the number of ASVs/OTUs is given above, while the fractions of reads for each amplicon are shown below.
Long amplicons were sequenced on PacBio RS II, short amplicons were sequenced on PacBio RS II, Illumina MiSeq, and Ion Torrent Ion S5.
ASV/OTU counts for short amplicons reflect detection by any of the three technologies, and read counts represent the mean fraction of reads across the three technologies.

```{r venn3, fig.cap = "(ref:venn2)", out.height = "30%", fig.show="hold"}
readd(asv_table) %>%
  mutate_if(is.numeric, ~ . / sum(.)) %>%
  mutate(long = pb_500,
         short = rowMeans(.[,c("pb_483", "SH-2257", "is_057")])) %>%
  select(seq, long, short) %>%
venndata(ASVs, cols = c("long", "short")) %>%
  vennplot_data(ASVs) %>% {
  plot(
     select(., set, ASVs) %>% deframe() %>% replace_na(0) %>% eulerr::venn(),
     legend = list(
         labels = c("Long amplicon", "Short amplicon")
     ),
     quantities = .$label,
     main = "ASV",
     fills = c("white", "slateblue1"))
  }

readd(otu_table) %>%
  mutate_if(is.numeric, ~ . / sum(.)) %>%
  mutate(long = pb_500,
         short = rowMeans(.[,c("pb_483", "SH-2257", "is_057")])) %>%
  select(seq, long, short) %>%
venndata(OTUs, cols = c("long", "short")) %>%
  vennplot_data(OTUs) %>% {
  plot(
    select(., set, OTUs) %>% deframe() %>% replace_na(0) %>% eulerr::venn(),
    legend = FALSE,
    labels = FALSE,
    quantities = .$label,
    main = "OTU",
     fills = c("white", "slateblue1")
  )
}
```

@@

Almost all of the short amplicons from all three technologies were between 240 and 375 bp long (Supplementary Figure \@ref(fig:full-length)a).
Although the length profile of the three requencing runs were very similar, Illumina MiSeq had the largest fraction of reads near the top of the range, followed by Ion Torrent Ion S5 and PacBio RS II (Supplementary Figure \@ref(fig:full-length)b).
In contrast, the length of the long amplicon reads varied widely, from `r filter(readd(reads_table), region == "long") %$% range(length) %>% paste(collapse = " to ")` bp, with a median of

@@

### Autocorrelation

@@

```{r variofit, results="asis"}
readd(variog_data) %>%
  select(mantel_0, mantel_1, variofit_0 = variofit2, variofit_1 = variofitST2, metric, amplicon, tech, guild, algorithm) %>%
  pivot_longer(1:4, names_to = c(".value", "timelag"), names_sep = "_") %>%
  mutate(params = map(variofit, broom::tidy)) %>%
  unnest(params) %>%
  bind_cols(
    pmap_dfr(
      .,
      function(mantel, variofit, term, ...) {
        if (mantel) {
          tryCatch(broom::confint_tidy(object = variofit, parm = term) %>% as.list %>% inset2("term", term) %>% as.data.frame,
                   error = function(e) tibble(term = term, conf.low = NA_real_, conf.high = NA_real_))
        } else {
          tibble(term = term, conf.low = NA_real_, conf.high = NA_real_)
        }
      }) %>%
      select(-term)
  ) %>%
  filter(term %in% c("range", "timerange")) %>%
  mutate_at("conf.low", replace_na, 0) %>%
  mutate(value = ifelse(
    is.na(estimate) | is.na(conf.high),
    "-",
    sprintf("%.2g (%.2g–%.2g)", estimate, conf.low, conf.high)
  )) %>%
  filter(timelag == "0" | term == "timerange") %>%
  pivot_wider(names_from = "term", values_from = "value", id_cols = c("guild", "amplicon", "tech", "metric", "algorithm")) %>%
  select(Guild = guild, Amplicon = amplicon, Tech = tech, Algorithm = algorithm, Metric = metric, "Space Range (m)" = range, "Time Range (a)" = timerange) %>%
  arrange(Guild, Amplicon, Tech, Algorithm, Metric) %>%
  kable(booktabs = TRUE, caption = "(ref:variofit)", linesep = c(rep("", 5), "\\addlinespace"))
```

@@

```{r phylotax, fig.height = 2, fig.width = 3, fig.cap = "(ref:phylotax)"}
readd(phylotax_fig)
```

(ref:full-length) Comparison of the length of denoised sequences from different length amplicons and sequencing technologies.
Length distribution (*a*, *c*) and empirical cumulative distribution function (ECDF; *b*, *d*) for short (*a*, *b*) and long (*c*, *d*) amplicons, respectively.

```{r full-length, fig.cap = "(ref:full-length)", fig.height = 6}
ggpubr::ggarrange(
  readd(full_dens_short),
  readd(full_ecdf_short),
  readd(full_dens_long),
  readd(full_ecdf_long),
  ncol = 1, heights = c(3, 1.7, 1.2, 2), labels = "auto", vjust = 1
)
```

@@

(ref:region-asv) Plot of total number of ASVs recovered vs. fraction of reads successfully mapped to ASVs for different rDNA regions extracted from long PacBio amplicons using `LSUx`. Some regions were formed by concatenation of extracted regions, as follows: LSU = LSU1--D1--LSU2--D2--LSU3--D3--LSU4; ITS = ITS1--5.8S--ITS2; 32S = 5.8S--ITS2--LSU; long = ITS--LSU. For this figure, subregions were concatenated prior to generating ASVs with `DADA2`.

```{r regions, fig.cap="(ref:region-asv)"}
readd(region_table) %>%
  dplyr::filter(seq_run == "pb_500") %>%
  mutate(reads = reads / sum(filter(readd(readcounts), seq_run == "pb_500", is.na(step), !is.na(well))$nreads)) %>%
  dplyr::mutate_at("region", fct_recode, "5.8S" = "5_8S") %>%
  ggplot(aes(reads, ASVs, color = length_med, label = region)) +
  geom_point() +
  ggrepel::geom_text_repel() +
  scale_color_gradientn(colors = c("blue", "red", "green1"), limits = c(30, 2000), breaks = c(50, 150, 400, 1000), trans = "log10", name = "median length") +
  scale_x_continuous(name = "Fraction of reads mapped",limits = c(0, 1))
```

@@

```{r region-lengths, fig.cap = "(ref:region-length)", fig.height = 6}
readd(region_length_fig)
```

@@

```{r ITS2-length, fig.cap = "(ref:ITS2-length)", fig.height = 6}
ggpubr::ggarrange(
  readd(its2_dens),
  readd(its2_ecdf),
  ncol = 1,
  labels = "auto",
  heights = c(4, 2)
)
```

@@

```{r short-its, fig.cap = "(ref:short-its)"}
readd(big_seq_table_ITS2) %>%
  colSums() %>%
  enframe(name = "ITS2", value = "nreads") %>%
  mutate(length = nchar(ITS2)) %>%
  filter(length <= 140) %>%
  mutate_at("ITS2", ~tzara::seqhash(chartr("T", "U", .))) %>%
  semi_join(readd(taxon_reads), ., by = c("label" = "ITS2")) %>%
  # group_by(ITS2) %>%
  # filter(!"Short" %in% amplicon) %>%
  taxon_plot(kingdom, y = reads, datasets = datasets)
```

@@

```{r ECM, fig.cap = "(ref:ecm)"}
  readd(taxon_reads) %>%
  mutate_at("reference", replace_na, "All") %>%
  taxon_plot(ECM, kingdom == "Fungi", datasets = datasets) +
  theme(legend.position = "bottom", legend.title = element_blank())
```
